Comparing probabilistic predictive models applied to football

### 2015-05-09 template.tex
alguns errinhos a menos; figura atualizada

\begin{table}[H]
\begin{center}
\begin{tabular}{cccccc}

\hline
Matchday & BB1 & BB2 & $Mn-Dir_1$ & $Mn-Dir_2$ & $Mn-Dir_3$\\
\hline
\hline
1 &4.992 & 5.055 & 5.630 & 5.053 & 5.521 \\
2 &5.006 & 5.328 & 5.327 & 5.533 & 5.201 \\
3 &7.129 & 6.568 & 7.979 & 7.640 & 8.216 \\
4 &5.941 & 5.782 & 6.513 & 6.848 & 6.564 \\
5 &5.246 & 5.858 & 5.669 & 5.368 & 5.601 \\
\hline
Total & 111.04 & 111.32 & 115.45 & 115.36 & 115.29 \\
\hline
Mean & 0.5844 & 0.5859 & 0.6076 & 0.6072 & 0.6068 \\
\hline
\end{tabular}
\caption{Score of the five first matchdays and mean score per match after 190 matches}
\end{center}
\end{table}

@@

Figure \ref{fig::scores} displays the boxplots of the scores and proportion of errors of the predictive models considered.
It also shows an horizontal line with the score of the trivial prediction $(1/3,1/3,1/3)$. 
All methods have similar performance, and they are more accurate than the trivial prediction, displayed in the plots
as an horizontal line.\footnote{None of the 95\% confidence intervals for the mean score contain the score given by the trivial prediction.}

Figure \ref{fig::entropy} presents the boxplots of the entropy of each prediction for the predictive models considered.\footnote{Recall
that the entropy of a prediction $(p_1,p_2,p_3)$ is given by $- \sum_{i=1}^3 p_i \log{p_i}$} It shows that the predictions
are typically more informative than $(1/3,1/3,1/3)$, but overall all methods yield similar entropies in average.

We evaluate how reasonable \added{were the?} predictions by assessing the calibration of the methods considered, i.e., by evaluating how often events which have assigned probability $p$ (for each $0<p<1$) happen \added{or happened?} (Section \ref{sec::scoring}).
If these probabilities are close to $p$, one concludes that the methods are well-calibrated. 
Results are displayed in Figure \ref{fig::calibration}.
Because we typically do not have several predictions with the same assigned probability $p$, we obtain the plot by smoothing the indicator function of whether a given event happened on  the assigned  probability of that event. 
The smoothing is done via smoothing splines \cite{wahba}, with tuning parameters chosen by cross-validation.
Because the curves obtain \added{(ou we obtained?)} are close to the identity, we conclude that all methods are well-calibrated, except perhaps BB1, in which events with both low and high estimated probabilities happen more often 
than predicted.

@@

@@

@@

@@

This work thus indicates that while there has been a boom in the literature with the proposal of several fancy and involved models, especially in the era of big data, simple models can many times do as well as their complex counterparts. 
This is often unnoticed by users of such methods, which leads to the overuse of models that are hard to interpret and to implement, and that also require one to keep track of several covariates.
We would call this phenomenon the {\it king's nudity}\footnote{In allusion to Anderson's {\it The Emperor's New Clothes}, \cite{emperor}: ``But he isn't wearing anything at all!''.}, that is not, of course, always observed.
In this spirit, since we have explored only the prediction of football matches, we would like to study if the the same extends to other fields such as, for instance, \red{exemplos} \red{Tentei aqui Rafael, veja o que acha.}

If, on the one hand, our findings appear as another supportive example of Ockham's razor, on the other they pose several questions about probabilistic prediction of sport events.
In particular, based on the fact that the models have similar performance, one may ask: is there an irreducible ``randomness'' or ``degree of unpredictability'' implicit in these events?
Is this degree an indicator of how tight or leveled is the championship being studied? \red{ficou mais clara a segunda pergunta? Queria perguntar se a dificuldade de prever os resultados indica o quao disputado ou equilibrado \'e o campeonato.}

>>>>

@@

Figure \ref{fig::scores} displays the boxplots of the scores and proportion of errors of the predictive models considered.
It also shows an horizontal line with the score of the trivial prediction $(1/3,1/3,1/3)$. 
According to all scoring rules, all methods have similar performance, and they are more accurate than the trivial prediction, displayed in the plots
as an horizontal line.\footnote{None of the 95\% confidence intervals for the mean score contain the score given by the trivial prediction.} The Brier score the first five matchdays, as well as the total Brier score is also displayed 
in Table \ref{tab::brier}.

@@

\begin{table}[H]
\begin{center}
\begin{tabular}{cccccc}
\hline
Matchday & BB1 & BB2 & $Mn-Dir_1$ & $Mn-Dir_2$ & $Mn-Dir_3$\\
\hline
\hline
1 &4.992 & 5.055 & 5.630 & 5.053 & 5.521 \\
2 &5.006 & 5.328 & 5.327 & 5.533 & 5.201 \\
3 &7.129 & 6.568 & 7.979 & 7.640 & 8.216 \\
4 &5.941 & 5.782 & 6.513 & 6.848 & 6.564 \\
5 &5.246 & 5.858 & 5.669 & 5.368 & 5.601 \\
\hline
Total & 111.04 & 111.32 & 115.45 & 115.36 & 115.29 \\
\hline
Mean & 0.5844 & 0.5859 & 0.6076 & 0.6072 & 0.6068 \\
\hline
\end{tabular}
\caption{Brier score of the five first matchdays and mean score per match after 190 matches.
The score of the trivial prediction $(1/3,1/3,1/3)$ is 0.667.}
\label{tab::brier}
\end{center}
\end{table}

Figure \ref{fig::entropy} presents the boxplots of the entropy of each prediction for the predictive models considered.\footnote{Recall
that the entropy of a prediction $(p_1,p_2,p_3)$ is given by $- \sum_{i=1}^3 p_i \log{p_i}$} 
Because the entropy values are smaller than 1.09, 
the plots shows that the predictions
are typically more informative than $(1/3,1/3,1/3)$. Nevertheless, all methods yield similar entropies in average, that is,
none of them seem to give more informative probabilities.

@@

We evaluate how reasonable were the predictions by assessing the calibration of the methods considered, i.e., by evaluating how often events which have assigned probability $p$ (for each $0<p<1$) happened (Section \ref{sec::scoring}).
If these probabilities are close to $p$, one concludes that the methods are well-calibrated. 
Results are displayed in Figure \ref{fig::calibration}.
Because we typically do not have several predictions with the same assigned probability $p$, we obtain the plot by smoothing the indicator function of whether a given event happened on  the assigned  probability of that event. 
The smoothing is done via smoothing splines \cite{wahba}, with tuning parameters chosen by cross-validation.
Because the curves we obtained are close to the identity, we conclude that all methods are well-calibrated, except perhaps BB1, in which events with both low and high estimated probabilities happen more often 
than predicted.

@@

We therefore conclude that all methods have similar predictive power, and none of them is more informative in average.
Nevertheless, all methods yield better predictions than the non-informative vector of probabilities $(1/3,1/3,1/3)$, and they are well-calibrated.

@@

This work thus indicates that while there has been a boom in the literature with the proposal of several fancy and involved models, especially in the era of big data, simple models can many times do as well as their complex counterparts. 
This is often unnoticed by users of such methods, which leads to the overuse of models that are hard to interpret and to implement, and that also require one to keep track of several covariates.
We would call this phenomenon the {\it king's nudity}\footnote{In allusion to Anderson's {\it The Emperor's New Clothes}, \cite{emperor}: ``But he isn't wearing anything at all!''.}, that is not, of course, always observed.
In this spirit, since we have explored only the prediction of football matches, we would like to study if the the same extends to other fields such as, for instance, \red{exemplos}

If, on the one hand, our findings appear as another supportive example of Ockham's razor, on the other they pose several questions about probabilistic prediction of sport events.
In particular, based on the fact that the models have similar performance, one may ask: is there an irreducible ``randomness'' or ``degree of unpredictability'' implicit in these events?
Is this degree an indicator of how tight or leveled is the championship being studied?
